Carbonate chemistry at Ny-Ålesund

### 2021-03-19 awipev-CO2.Rmd
revised choice of constants with discrete samples

# calculate means
monthly_means <- d_all %>% 
  group_by(montht) %>% 
  summarize(mean_ct = mean(ct, na.rm = TRUE),
            mean_at_calc = mean(at_calc, na.rm = TRUE),
            mean_pco2_calc = mean(pco2_calc_fb, na.rm = TRUE),
            mean_sal = mean(sal_fb, na.rm = TRUE),
            mean_temp = mean(temp_fb, na.rm = TRUE)) %>% 
  ungroup()
means <- monthly_means %>% 
  summarise(across(
    .cols = is.numeric, 
    .fns = mean, na.rm = TRUE, 
    .names = "{col}"
  ))

mean_ct <-  mean(monthly_means$mean_ct)
mean_at_calc <-  mean(monthly_means$mean_at_calc)
mean_pco2_calc <-  mean(monthly_means$mean_pco2_calc)
mean_sal <-  mean(monthly_means$mean_sal)
mean_temp <-  mean(monthly_means$mean_temp)
# impacts using pco2 and AT pair
carb_l_1 <- carb(24, mean_pco2_calc, mean_at_calc*1e-6, S = mean_sal, T = mean_temp, P = 0, 
                 Pt = 0, Sit = 0,
                 k1k2 = "l", kf = "dg", ks = "d", pHscale = "T", b = "u74")
carb_s_1 <- carb(24, mean_pco2_calc, mean_at_calc*1e-6, S = mean_sal, T = mean_temp, P = 0, 
                 Pt = 0, Sit = 0,
                 k1k2 = "s20", kf = "dg", ks = "d", pHscale = "T", b = "u74")
diff_ls_1 <- carb_l_1 - carb_s_1

# impacts using AT-DIC pair
carb_l_2 <- carb(15, mean_at_calc*1e-6, mean_ct*1e-6, S = mean_sal, T = mean_temp, P = 0, 
                 Pt = 0, Sit = 0,
                 k1k2 = "l", kf = "dg", ks = "d", pHscale = "T", b = "u74")
carb_s_2 <- carb(15, mean_at_calc*1e-6, mean_ct*1e-6, S = mean_sal, T = mean_temp, P = 0,
                Pt = 0, Sit = 0, 
                k1k2 = "s20", kf = "dg", ks = "d", pHscale = "T", b = "u74")
diff_ls_2 <- carb_l_2 - carb_s_2
const_choice <- bind_rows(diff_ls_1, diff_ls_2) %>% 
  dplyr::mutate(flag = c("pCO2 and ALK", "ALK and DIC"),
                S = mean_sal,
                T = mean_temp,
                DIC = DIC * 1e6
                ) %>% 
  dplyr::select(flag, pH, pCO2, DIC, OmegaAragonite)
```
The means used are:
`r knitr::kable(means, format = "simple")`

The differences are:
`r knitr::kable(const_choice, digits = c(0, 3, 1, 1, 3), caption = "Lueker minus Sulpis", format = "simple")`

### Lueker vs Sulpis using all values pCO2 and at_calc
One uses the pair pCO2-AT_calc

```{r impact of constants 2, echo=FALSE, warning=FALSE, message=FALSE}
# Compare Lueker and Sulpis using pco2_fb and at_calc
tmp <- d_all %>% 
  tidyr::drop_na(pco2_fb, at_calc, sal_fb, temp_fb) %>% 
  dplyr::select(ph_s_sf_norm4, pco2_fb, at_calc, sal_fb, temp_fb)

@@

# calculate quantiles
diff_ls_quant <- bind_rows(
  quantile(diff_ls$delta_ph), quantile(diff_ls$delta_ct * 1e6),
  quantile(diff_ls$delta_OmegaAragonite)) %>% 
  dplyr::mutate(Variable = c("pH", "DIC", "OmegaAragonite"), .before = 1)
#pander(diff_ls_quant)

@@

p1 <- ggreg(xdata = tmp2$ph_s_sf_norm4, ydata = tmp2$pH_l, type = "OLS", plot = TRUE, 
            title = "both pH normalized at 4°C", subtitle = "",
            xlab = "Spectrophotometric pH measured @ 4°C",
            ylab = "pH @ 4°C calculated (pCO2+AT) with Lueker")
p2 <- ggreg(xdata = tmp2$ph_s_sf_norm4, ydata = tmp2$pH_s, type = "OLS", plot = TRUE, title = "both pH normalized at 4°C")

lm <- lm2(x = tmp2$ph_s_sf_norm4, y = tmp2$pH_l)
p <-  ggplot(d_all,aes(x=sal_insitu_ctd_9m, y= sal_fb))+
  geom_point( aes(color=factor(monthd))) + 
  geom_abline(slope=1, intercept=0, linetype = "dashed")+
  geom_abline(aes(intercept = lm$fit$regression.results[2,2], slope = lm$fit$regression.results[2,3]),colour = "blue") +
  ylim(c(30.5,35.5)) +xlim(c(28,35.5)) +
  labs(title = paste("","\n Adj R2 = ", signif(lm$fit$rsquare, 3),
                     "; Intercept =", signif(lm$fit$regression.results[2,2], 3),
                     "; Slope =", signif(lm$fit$regression.results[2,3], 3),
                     "; P =", signif(lm$fit$P.param, 3),
                     "; RMSE = ", signif(lm$rmse, 3))) +
  theme(aspect.ratio = 1, plot.title = element_text(size = 7)) + #
  coord_fixed(ratio = 1 )+
  labs( x="Salinity CTD 9 m", y="Salinity FB ", color = "Months") + theme_bw()

The difference induced by the choice of the carbonic acid constants is large for pH, pCO2 and DIC and negligible for OmegaAragonite.

@@

>>>>

# # calculate means
# monthly_means <- d_all %>% 
#   group_by(montht) %>% 
#   summarize(mean_ct = mean(ct, na.rm = TRUE),
#             mean_at_calc = mean(at_calc, na.rm = TRUE),
#             mean_pco2_calc = mean(pco2_calc_fb, na.rm = TRUE),
#             mean_sal = mean(sal_fb, na.rm = TRUE),
#             mean_temp = mean(temp_fb, na.rm = TRUE)) %>% 
#   ungroup()
# means <- monthly_means %>% 
#   summarise(across(
#     .cols = is.numeric, 
#     .fns = mean, na.rm = TRUE, 
#     .names = "{col}"
#   ))

mean_ct <-  mean(discrete$ct, na.rm = TRUE)
mean_at <-  mean(discrete$ct, na.rm = TRUE)
mean_pco2 <-  mean(d_all$pco2_calc_fb)
mean_sal <-  mean(discrete$sal_fb, na.rm = TRUE)
mean_temp <-  mean(discrete$temp_fb, na.rm = TRUE)

# calculate typical uncertainty assuming uncertainties of 3 uatm for pCO2 and 4 umol/kg for AT
errors_pco2_at <- seacarb::errors(flag=24, var1=mean_pco2_calc, var2=1e-6*mean_at_calc, S=mean_sal,
                          T=mean_temp, P=0, Pt=0, Sit=0, k1k2="l", kf="dg", ks="d", pHscale="T",
                          b="u74", evar1=3, evar2 = 4e-6)
errors_at_ct <- seacarb::errors(flag=15, var1=1e-6*mean_at_calc, var2=1e-6*mean_ct, S=mean_sal,
                          T=mean_temp, P=0, Pt=0, Sit=0, k1k2="l", kf="dg", ks="d", pHscale="T",
                          b="u74", evar1=4e-6, evar2 = 4e-6)
errors <- tibble(e_pH = c(errors_pco2_at$pH, errors_at_ct$pH),
                 e_pCO2 = c(NA, errors_at_ct$pCO2),
                 e_DIC = c(errors_pco2_at$pH, NA)*1e3,
                 e_Oa = c(errors_pco2_at$OmegaAragonite, errors_at_ct$OmegaAragonite)
)

# impacts using pco2 and AT pair
carb_l_1 <- carb(24, mean_pco2_calc, mean_at_calc*1e-6, S = mean_sal, T = mean_temp, P = 0, 
                 Pt = 0, Sit = 0,
                 k1k2 = "l", kf = "dg", ks = "d", pHscale = "T", b = "u74")
carb_s_1 <- carb(24, mean_pco2_calc, mean_at_calc*1e-6, S = mean_sal, T = mean_temp, P = 0, 
                 Pt = 0, Sit = 0,
                 k1k2 = "s20", kf = "dg", ks = "d", pHscale = "T", b = "u74")
diff_ls_1 <- carb_l_1 - carb_s_1

# impacts using AT-DIC pair
carb_l_2 <- carb(15, mean_at_calc*1e-6, mean_ct*1e-6, S = mean_sal, T = mean_temp, P = 0, 
                 Pt = 0, Sit = 0,
                 k1k2 = "l", kf = "dg", ks = "d", pHscale = "T", b = "u74")
carb_s_2 <- carb(15, mean_at_calc*1e-6, mean_ct*1e-6, S = mean_sal, T = mean_temp, P = 0,
                Pt = 0, Sit = 0, 
                k1k2 = "s20", kf = "dg", ks = "d", pHscale = "T", b = "u74")
diff_ls_2 <- carb_l_2 - carb_s_2
const_choice <- bind_rows(diff_ls_1, diff_ls_2) %>% 
  dplyr::mutate(flag = c("pCO2 and ALK", "ALK and DIC"),
                S = mean_sal,
                T = mean_temp,
                DIC = DIC * 1e6
                ) %>% 
  dplyr::select(flag, pH, pCO2, DIC, OmegaAragonite) %>% 
  dplyr::bind_cols(errors)
```
The means used are:
`r knitr::kable(means, format = "simple")`

The differences are:
`r knitr::kable(const_choice, digits = c(0, 3, 1, 1, 3, 3, 1, 1, 1), caption = "Lueker minus Sulpis. e indicates typical errors estimated with seacarb.", format = "simple")`

### Lueker vs Sulpis using all values pCO2 and at_calc
```{r impact of constants 2, echo=FALSE, warning=FALSE, message=FALSE}
# Compare Lueker and Sulpis using pco2_fb and at_calc
tmp <- d_all %>% 
  tidyr::drop_na(pco2_fb, at_calc, sal_fb, temp_fb) %>% 
  dplyr::select(ph_s_sf_norm4, pco2_fb, at_calc, sal_fb, temp_fb)

@@

# calculate descriptive statistics
diff_ls_quant <- tibble(pH = as.numeric(summary(diff_ls$delta_ph)),
                        CT = as.numeric(summary(diff_ls$delta_ct * 1e6)),
                        OmegaAragonite = as.numeric(summary(diff_ls$delta_OmegaAragonite))) %>%
  dplyr::mutate(var = c("Min", "Q1", "Median", "Mean", "Q3", "Max"), .before = 1)

@@

# summary of differences
diff_lueker <- as.numeric(summary(tmp2$ph_s_sf_norm4 - tmp2$pH_l))
diff_sulpis <- as.numeric(summary(tmp2$ph_s_sf_norm4 - tmp2$pH_s))
lueker_sulpis_table <- tibble(var = c("Min", "Q1", "Median", "Mean", "Q3", "Max"), diff_lueker, diff_sulpis)

# make plots
p1 <- ggreg(xdata = tmp2$ph_s_sf_norm4, ydata = tmp2$pH_l, type = "MA", plot = TRUE, 
            title = "Lueker vs measured", subtitle = "both pH normalized at 4°C",
            xlab = "Spectrophotometric pH measured @ 4°C",
            ylab = "pH @ 4°C calculated (pCO2+AT) with Lueker")
p2 <- ggreg(xdata = tmp2$ph_s_sf_norm4, ydata = tmp2$pH_s, type = "MA", plot = TRUE, 
            title = "Sulpis vs measured", subtitle = "both pH normalized at 4°C",
            xlab = "Spectrophotometric pH measured @ 4°C",
            ylab = "pH @ 4°C calculated (pCO2+AT) with Sulpis") 
p <- p1 + p2
ggsave(filename = "figures/lueker-sulpis.png", plot = p, width = 20, height = 15, units = "cm")
```
The overall means of pCO2, AT (calculated from the AT-salinity relationship), salinity and temperature in the FerryBox were used to calculate pH from measured pCO2 and AT (from a salinity-AT relationship) and from AT and CT. using the formulations of Lueker et al. (2000) and Sulpis et al. (2020) of K1 and K2. The difference induced by the choice of the carbonic acid constants is large for pH, pCO2 and DIC and negligible for OmegaAragonite.

@@

Here we compare pH measured spectrophotometrically with pH calculated from pCO2 and AT using the sets of K1 and K2 of Lueker et al. (2000) and Sulpis et al. (2020). All pH values were normalized to a temperature of 4°C. Both formulations overestimate spectrophotometric pH 
`r knitr::kable(lueker_sulpis_table, digits = 3, caption = "Difference between spectrophotometric pH and Lueker and Sulpis", format = "simple")`
